### 2022-02-16 misc/solutions/prob-method-sols.tex
Did 4.4

>>>>

\proclaim Exercise 4.4. Let $X$ be a random variable with expectation $\ex\{X\} =0$ and variance
$\var\{X\} = \sigma^2$. Prove that for all $\lambda > 0$,
$$\pr\{X\ge \lambda\} \le {\sigma^2\over \sigma^2 + \lambda^2}.$$

\proof Let $\lambda > 0$. Note that if $X\ge \lambda$, then $X$ is also positive so in particular,
for any $a>0$, $X+a$ and $\lambda + a$ are both positive. This implies that $(X+a)^2\ge (\lambda+a)^2$.
Note that since $\var\{X\} = \ex\{X^2\} - \ex\{X\}^2 = \sigma^2$ and $\ex\{X\}=0$, we have
$\ex\{X^2\} = \sigma^2$, which in turn implies that
$$\ex\bigl\{(X+a)^2\bigr\} = \ex\{X^2\} + 2a\ex\{X\} + \ex\{a^2\} = \sigma^2+a^2.$$
Now we take the infimum over all $a>0$ and apply Markov's inequality to get
$$\pr\{X\ge\lambda\} \le \inf_{a > 0} \pr\bigl\{(X+a)^2 \ge (\lambda+a)^2\bigr\} \le
\inf_{a>0} {\sigma^2 + a^2\over (\lambda + a)^2}.$$
We now optimise over $a>0$. We have
$${d\over da} {\sigma^2 + a^2\over (\lambda + a)^2}
  = {(\lambda+a)^2 2a - (\sigma^2+a^2)2(\lambda+a)\over (\lambda + a)^4},$$
which is zero when $a = \sigma^2/\lambda > 0$. Plugging in this value of $a$ gives
$$\pr\{X\ge \lambda\} \le {\sigma^2 + \sigma^4/\lambda^2\over (\lambda + \sigma^2/\lambda)^2}
= {\sigma^2(\lambda + \sigma^2/\lambda) \over \lambda(\lambda + \sigma^2/\lambda)^2}
= {\sigma^2 \over \sigma^2+\lambda^2},$$
exactly what we need.\slug
